---
title: "Tutorial_script"
author: "Lucas Mallet"
date: "`r Sys.Date()`"
output: html_document
---

#  *Operators*
## Arithmetic

| sign | description                                              | example |
| :--: | :------------------------------------------------------- | :-----: |
|  +   | **Addition** of two operators                            |  a + b  |
|  -   | **Subtraction** of second operand from first             |  a – b  |
|  **  | **Multiplication** of two operands                       |  a * b  |
|  /   | **Division** of first operand with second                |  a / b  |
|  %%  | **Remainder** from division of first operand with second | a %% b  |
| %/%  | **Quotient** from division of first operand with second  | a %/% b |
|  ^   | First operand raised to the **power** of second operand  |   a^b   |

## Relational

| sign | description                                                  | example |
| :--: | :----------------------------------------------------------- | :-----: |
|  <   | Is first operand **less** than second operand                |  a < b  |
|  >   | Is first operand **greater** than second operand             |  a > b  |
|  ==  | Is first operand **equal** to second operand                 | a == b  |
|  <=  | Is first operand **less than or equal** to second operand    | a <= b  |
|  >=  | Is first operand **greater than or equal** to second operand | a >= b  |
|  !=  | Is first operand **not equal** to second operand             | a != b  |

## Logical

| sign | description                            | example  |
| :--: | :------------------------------------- | :------: |
|  &   | Element wise logical **AND** operation |  a & b   |
|  \|  | Element wise logical **OR** operation  |  a \| b  |
|  !   | Element wise logical **NOT** operation |    !a    |
|  &&  | Operand wise logical **AND** operation |  a && b  |
| \|\| | Operand wise logical **OR** operation  | a \|\| b |

## Assignement
|  sign  | description                                   | example |
| :----: | :-------------------------------------------- | :-----: |
| = / <- | Assigns right side value to left side operand | a <- 3  |
|   ->   | Assigns left side value to right side operand | 4 -> b  |

## Miscellaneous

| sign | description                                                                                   |  example   |
| :--: | :-------------------------------------------------------------------------------------------- | :--------: |
|  :   | Creates series of numbers from left operand to right operand                                  |    a:b     |
| %in% | Identifies if an element "a" belongs to a vector "b"                                          |  a %in% b  |
| %*%  | Performs multiplication of a vector with its transpose                                        | A %*% t(A) |
| %>%  | Creates a pipe where the first object is integrated in the second (alternative to nestedness) | a %>% b()  |

## Other

|  sign  | description                                                                                                                                                                                                                                                                                                                                                           |
| :----: | :-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
|  ( )   | Round brackets (also known as "parenthesis") are used primarily when calling a function in R. Every function must be called using the round brackets.                                                                                                                                                                                                                 |
|  [ ]   | The square brackets are used for indexing into a vector, matrix, array, list or dataframe. You can think of the square brackets as marking the edges of a cell, column or row of a table. The square brackets are also called _extraction operators_ as they are used to _extract_ specific elements from a vector or matrix. _["number of row", "number of column"]_ |
| \[[ ]] | allows only a single element to be selected using integer or character indices, whereas `[` allows indexing by vectors.                                                                                                                                                                                                                                               |
|  { }   | Curly braces are used to group multiple expressions together into a **block of code**. They ensure that the grouped expressions are executed as a single unit                                                                                                                                                                                                         |

# *Objects* 
## Basic

| type of object | description                                                                                                                                                                                                                                         |
| :------------: | --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
|    Numeric     | -  **Double** : allows you to store numbers as decimals. This is the default treatment for numbers<br>- **Integer** : an integer is a subset of the numeric data type. This type will only allow whole numbers and is denoted by the letter “L”<br> |
|    Complex     | imaginary number through the use of the lowercase letter “i”                                                                                                                                                                                        |
|   Character    | store text data : between quote as "text"                                                                                                                                                                                                           |
|    Logical     | boolean data : TRUE / FALSE                                                                                                                                                                                                                         |
## Complex
| type of object | description                                                                                                                                                                                                                                |
| :------------: | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ |
|     Vector     | ordered collection of numbers or character strings indexed by the integers 1, 2, …, n, where n is the length of the vector.                                                                                                                |
|     Factor     | special kind of vector which contains underlying numeric values 1, 2, …, n, but each of these n values has an associated character label (which may or may not be the numeric value). These labeled values are the **levels** of the facto |
|     Matrix     | two-dimensional collection of numeric or character values indexed by pairs of integers (i, j).                                                                                                                                             |
|     Array      | n-dimensional collection of numeric or character values indexed by n-tuples of integers (e.g., (i,j,k) for a 3-dimensional array                                                                                                           |
|      List      | an ordered collection of objects, and the objects in the list can be of different types.                                                                                                                                                   |
|   Data frame   | type of list commonly used to store datasets, with a row for each observation and a column for each variable, and each variable can be of a different type                                                                                 |

# *Examples* (R script)
## Operator 
```{r}
# Arithmetic
x <- 5
y <- 16
x + y
x - y
x * y
y / x
y %/% x
y %% x
y ^ x

# Relational 
x < y
x > y
x <= 5
y >= 20
y == 16
x != 5

# Logical
a <- c(FALSE,  0,     6,     TRUE) ## Zero is considered FALSE and non-zero numbers are taken as TRUE. Let's see an example for this:
b <- c(TRUE,   FALSE, TRUE,  TRUE)
!a
a & b
b & a
a && b
a | b
a || b 

# Miscellaneous 
library("dplyr")

## :
u <- c(2:9) 

## %in%
c("1","4", "2.8") %in% u

## %*%
x <- 1:4
(z <- x %*% x)    # scalar ("inner") product as : ∑(xi.yi) <=> z=1*1+2*2+3*3+4*4 = 30
drop(z)             # drop() delete tge dimension of array which have only one level. Becomes scalar

(y <- diag(x))
(z <- matrix(1:12, ncol = 3, nrow = 4)) 
y %*% z
y %*% x
x %*% z

## %>%
v <- c(1.2, 8.7, 5.6, 11.4)
v %>% 
  mean() %>%
  round(digits = 1) %>%
  format(decimal.mark = ",") %>%
  paste0("La moyenne est de ", .,".")

paste0("La moyenne est de ",(format(round(mean(v), digits = 1), decimal.mark = ",")),".") # with parenthesis : struggle for reading code
```

## Specific to curly braces
###  Defining block codes 
When curly braces are used, all the expressions inside them are evaluated together. The result of the last evaluated expression is returned. For example:
```{r}
{
  x <- 10
  y <- 20
  x + y
}
```

### Control flow statements
Curly braces are required in control flow structures like `if`, `for`, and `while` to group the statements executed within the block.
``` {r}
x <- 5
if (x > 3) {
  print("x is greater than 3")
}

```

### Functions definition
Curly braces define the body of a function in R. They contain the statements that the function will execute.
``` {r}
my_function <- function(a, b) {
  result <- a + b
  return(result)
}
my_function(3, 4)  
```

### Anonymous blocks for clarity
``` {r}
{
  temp <- sqrt(25)
  print(temp)
}
```

### Scoping
Curly braces can be used to limit the scope of variables. Variables declared inside a `{}` block are not accessible outside that block.
``` {r}
local({
  y <- 10  
  print(y) 
})
print(y) 
```

## Object
```{r}
# numeric
x <- 5
x
class(x)

# character
y <- "elevated blood pressure"
y
class(y)

# vector
z <- c(5, 8, 12)
z 
z[2] # second element of vector

# factor
y <- factor(c("poulet", "poule", "poule", "coq"),
            levels = c("poulet", "poule", "coq"),
            labels = c("POT", "POL", "COQ"))
levels(y)

# matrix
x <- matrix(c(1,2,3,4,5,6), nrow = 2, ncol = 3)
x

# array
z <- array(c(1,2,3,4,5,6,7,8), dim = c(2,2,2))
z

# list
x <- list("5", c(1,2,3), y) # I put y, an object we created earlier, in this list
x

x[[2]]
x[[3]]

# dataframe
d <- data.frame(outcome = c(1,0,1,1),
                exposure = c("yes", "yes", "no", "no"),
                age = c(24, 55, 39, 18))
d
```

