---
title: "Tutorial_script"
author: "Lucas Mallet"
date: "`r Sys.Date()`"
output: html_document
---

###  Operators
#### Arithmetic

| sign | description                                              | example |
| :--: | :------------------------------------------------------- | :-----: |
|  +   | **Addition** of two operators                            |  a + b  |
|  -   | **Subtraction** of second operand from first             |  a – b  |
|  **  | **Multiplication** of two operands                       |  a * b  |
|  /   | **Division** of first operand with second                |  a / b  |
|  %%  | **Remainder** from division of first operand with second | a %% b  |
| %/%  | **Quotient** from division of first operand with second  | a %/% b |
|  ^   | First operand raised to the **power** of second operand  |   a^b   |
#### Relational

| sign | description                                                  | example |
| :--: | :----------------------------------------------------------- | :-----: |
|  <   | Is first operand **less** than second operand                |  a < b  |
|  >   | Is first operand **greater** than second operand             |  a > b  |
|  ==  | Is first operand **equal** to second operand                 | a == b  |
|  <=  | Is first operand **less than or equal** to second operand    | a <= b  |
|  >=  | Is first operand **greater than or equal** to second operand | a >= b  |
|  !=  | Is first operand **not equal** to second operand             | a != b  |
#### Logical

| sign | description                            | example  |
| :--: | :------------------------------------- | :------: |
|  &   | Element wise logical **AND** operation |  a & b   |
|  \|  | Element wise logical **OR** operation  |  a \| b  |
|  !   | Element wise logical **NOT** operation |    !a    |
|  &&  | Operand wise logical **AND** operation |  a && b  |
| \|\| | Operand wise logical **OR** operation  | a \|\| b |
#### Assignement
|  sign  | description                                   | example |
| :----: | :-------------------------------------------- | :-----: |
| = / <- | Assigns right side value to left side operand | a <- 3  |
|   ->   | Assigns left side value to right side operand | 4 -> b  |
#### Miscellaneous

| sign | description                                                                                   |  example   |
| :--: | :-------------------------------------------------------------------------------------------- | :--------: |
|  :   | Creates series of numbers from left operand to right operand                                  |    a:b     |
| %in% | Identifies if an element "a" belongs to a vector "b"                                          |  a %in% b  |
| %*%  | Performs multiplication of a vector with its transpose                                        | A %*% t(A) |
| %>%  | Creates a pipe where the first object is integrated in the second (alternative to nestedness) | a %>% b()  |
#### Other

|  sign  | description                                                                                                                                                                                                                                                                                                                                                           |
| :----: | :-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
|  ( )   | Round brackets (also known as "parenthesis") are used primarily when calling a function in R. Every function must be called using the round brackets.                                                                                                                                                                                                                 |
|  [ ]   | The square brackets are used for indexing into a vector, matrix, array, list or dataframe. You can think of the square brackets as marking the edges of a cell, column or row of a table. The square brackets are also called _extraction operators_ as they are used to _extract_ specific elements from a vector or matrix. _["number of row", "number of column"]_ |
| \[[ ]] | allows only a single element to be selected using integer or character indices, whereas `[` allows indexing by vectors.                                                                                                                                                                                                                                               |
|  { }   | Curly braces are used to group multiple expressions together into a **block of code**. They ensure that the grouped expressions are executed as a single unit                                                                                                                                                                                                         |

### Objects 
#### Basic

| type of object | description                                                                                                                                                                                                                                         |
| :------------: | --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
|    Numeric     | -  **Double** : allows you to store numbers as decimals. This is the default treatment for numbers<br>- **Integer** : an integer is a subset of the numeric data type. This type will only allow whole numbers and is denoted by the letter “L”<br> |
|    Complex     | imaginary number through the use of the lowercase letter “i”                                                                                                                                                                                        |
|   Character    | store text data : between quote as "text"                                                                                                                                                                                                           |
|    Logical     | boolean data : TRUE / FALSE                                                                                                                                                                                                                         |
#### Complex
| type of object | description                                                                                                                                                                                                                                |
| :------------: | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ |
|     Vector     | ordered collection of numbers or character strings indexed by the integers 1, 2, …, n, where n is the length of the vector.                                                                                                                |
|     Factor     | special kind of vector which contains underlying numeric values 1, 2, …, n, but each of these n values has an associated character label (which may or may not be the numeric value). These labeled values are the **levels** of the facto |
|     Matrix     | two-dimensional collection of numeric or character values indexed by pairs of integers (i, j).                                                                                                                                             |
|     Array      | n-dimensional collection of numeric or character values indexed by n-tuples of integers (e.g., (i,j,k) for a 3-dimensional array                                                                                                           |
|      List      | an ordered collection of objects, and the objects in the list can be of different types.                                                                                                                                                   |
|   Data frame   | type of list commonly used to store datasets, with a row for each observation and a column for each variable, and each variable can be of a different type                                                                                 |

### Examples (R script)
##### Curly braces


######  defining block codes 
When curly braces are used, all the expressions inside them are evaluated together. The result of the last evaluated expression is returned. For example:
```{r}
{
  x <- 10
  y <- 20
  x + y
}
```

###### control flow statements
Curly braces are required in control flow structures like `if`, `for`, and `while` to group the statements executed within the block.
``` {r}
x <- 5
if (x > 3) {
  print("x is greater than 3")
}

```

###### functions definition
Curly braces define the body of a function in R. They contain the statements that the function will execute.
``` {r}
my_function <- function(a, b) {
  result <- a + b
  return(result)
}
my_function(3, 4)  
```

###### Anonymous blocks for clarity
``` {r}
{
  temp <- sqrt(25)
  print(temp)
}
```

###### scoping
Curly braces can be used to limit the scope of variables. Variables declared inside a `{}` block are not accessible outside that block.
``` {r}
local({
  y <- 10  
  print(y) 
})
print(y) 
```

